---
title: "Model Selection By AIC"
author: "Lukas"
date: "2023-03-28"
output: html_document
---

```{r}
library(nlme)
library(lme4)
library(VGAM)
library(car)
library(tseries)
library(forecast)
library(TSA)
library(doParallel)
library(parallel)
library(foreach)
library(vars)
library(astsa)
```


Load MATLAB Data
```{r}
Aug_08 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_08_table.csv")
Aug_09 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_09_table.csv")
Aug_12 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_12_table.csv")
Aug_15 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_15_table.csv")
Oct_12 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Oct_12_table.csv")
```

Remove Bad Tones from August 8th

```{r}
Aug_08 <- Aug_08[-c(1:60),]
```

Relabel Days
```{r}
Aug_08$Day <- 1
Aug_09$Day <- 2
Aug_12$Day <- 3
Aug_15$Day <- 4
Oct_12$Day <- 5
```



Day will likely be more useful when the days are all ran together with no gaps

Full Ephys Dataset
```{r}
Data <- rbind(Aug_08,Aug_09,Aug_12,Aug_15)
Test_Data <- Oct_12
rm(Aug_08,Aug_09,Aug_12,Aug_15,Oct_12)
```


Factors
```{r}
Data$Day <- as.factor(Data$Day)
Data$Tone <- as.factor(Data$Tone)
Data$Bin <- as.factor(Data$Bin)
```


Load Behavior Data
```{r}
## I Need to run August 5th through Analysis
Aug_08_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_08_new_R_TB_PMA17.csv")
Aug_09_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_09_new_R_TB_PMA17.csv")
Aug_12_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_12_new_R_TB_PMA17.csv")
Aug_15_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_15_new_R_TB_PMA17.csv")
Oct_12_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\10_12_new_R_TB_PMA17.csv")
```

Pull Time Specific Bins currently time 0 to 29. Will likely be adjusted to only tones

```{r}
Aug_08_Behav_p  <-  Aug_08_Behav[Aug_08_Behav[,2] %in% c(0:29),3]
Aug_09_Behav_p  <-  Aug_09_Behav[Aug_09_Behav[,2] %in% c(0:29),3]
Aug_12_Behav_p  <-  Aug_12_Behav[Aug_12_Behav[,2] %in% c(0:29),3]
Aug_15_Behav_p  <-  Aug_15_Behav[Aug_15_Behav[,2] %in% c(0:29),3]
Oct_12_Behav_p  <-  Oct_12_Behav[Oct_12_Behav[,2] %in% c(0:29),3]
```

All Behave

```{r}
#Remove same August Stuff
Aug_08_Behav_p <- Aug_08_Behav_p[-c(1:60)]
Behave <- c(Aug_08_Behav_p,Aug_09_Behav_p,Aug_12_Behav_p,Aug_15_Behav_p)

Test_Data$Platform <- Oct_12_Behav_p
Data$Platform <- Behave
```


Clean
```{r}
rm(Aug_08_Behav,Aug_09_Behav,Aug_12_Behav,Aug_15_Behav,Oct_12_Behav)
rm(Aug_08_Behav_p,Aug_09_Behav_p,Aug_12_Behav_p,Aug_15_Behav_p,Oct_12_Behav_p,Behave)
```

Data needs to be adjusted to accommodate 0 and 1

```{r}
#Playing it safe using a loop

offset <- 0.02


for(i in 1:length(Data$Platform)){
  
  if(Data$Platform[i] == 0){Data$Platform[i] <- offset}
  if(Data$Platform[i] == 1){Data$Platform[i] <- 1 - offset}
}

for(i in 1:length(Test_Data$Platform)){
  
  if(Test_Data$Platform[i] == 0){Test_Data$Platform[i] <- offset}
  if(Test_Data$Platform[i] == 1){Test_Data$Platform[i] <- 1 - offset}
}
```

Transform Response to make it fit regression requirement
```{r}
Data$probitPlatform <- probitlink(Data$Platform)
Test_Data$probitPlatform <- probitlink(Test_Data$Platform)
Test_Data.gls <- Test_Data
```

```{r}
Datats <- ts(Data$probitPlatform,frequency = 30,start = c(1,1))
model_var <-  ts(Data[,c(4:39)],frequency = 30,start = c(1,1))
Test_Data <- ts(Test_Data[,c(4:39,113)])
```

```{r}
start_model_means <- arimax(Datats,order=c(1,0,2),
                 seasonal=list(order=c(2,0,0),period=30))#,xreg=model_var[,18])
```


GLS
```{r}
gls.data <- Data[,c(1:39,113)]
gls.data$Day <- as.factor(gls.data$Day)
gls.data$Tone <- as.factor(gls.data$Tone)
gls.data$Bin <- as.factor(gls.data$Bin)
nesting.info <- groupedData(probitPlatform~1|Day/Tone/Bin,data=gls.data)

test.fit.gls <- gls(probitPlatform~Day+Tone+Bin,data=nesting.info,correlation = corARMA(,form=~1|Day/Tone,p=1,q=1),method = "ML")
```


Stepwise Longitudinal Model


```{r}
test_candidate_model_gls <- function(gls.var.available,i){
  candidate_model <- nlme::gls(as.formula(paste("probitPlatform~",paste(c(model_var_gls,gls.var.available[i]),collapse="+"))),data=nesting.info,correlation = nlme::corARMA(,form=~1|Day/Tone,p=1,q=1),method = "ML")
   
   test_aic <- extractAIC(candidate_model)[2]
   return(test_aic)
}
```




```{r}
model_var_gls <- c("Day","Tone","Bin")
gls.var <- names(nesting.info)[-c(1:3,40)]
gls.var.available <- gls.var

best_aic_gls <- extractAIC(test.fit.gls)[2]

check <-  F

cl <- parallel::makeCluster(14)
doParallel::registerDoParallel(cl)


while(check == F){
  
  candidates <- foreach::foreach(i = 1:length(gls.var.available), .combine = cbind) %dopar% {
    test_candidate_model_gls(gls.var.available = gls.var.available, i=i)
  }
  
  if(min(candidates) >= best_aic_gls){check = T}  
  
  if(min(candidates) < best_aic_gls & check == F){
    best_aic_gls <- min(candidates)
    model_var_gls <- c(model_var_gls,gls.var.available[sort(candidates,index.return=T)$ix[1]])
    gls.var.available <- gls.var.available[-sort(candidates,index.return=T)$ix[1]]
    }  
  print("Iteration")
}

parallel::stopCluster(cl)
model_var_gls

```

Final Model by AIC
```{r}
AIC_mod.gls <- gls(as.formula(paste("probitPlatform~",paste(model_var_gls,collapse="+"))),data=nesting.info,correlation = nlme::corARMA(c(0.7762,0.0972),form=~1|Day/Tone,p=1,q=1),method = "ML")

AIC_mod.gls.DayPoly <- gls(probitPlatform ~ poly(Day,2) +Tone + Bin + ILBLA_mean_High_Beta + PLBLA_mean_High_Beta + 
    IL_pow_mean_Low_Beta + BLA_pow_mean_Alpha + ILPL_mean_High_Beta + 
    ILBLA_mean_Mid_Beta + BLA_pow_mean_Mid_Beta + PL_pow_mean_High_Beta + 
    PL_pow_mean_Low_Beta + PL_pow_mean_Theta + PLBLA_mean_Delta,data=nesting.info,correlation = nlme::corARMA(c(0.7762,0.0972),form=~1|Day/Tone,p=1,q=1),method = "ML")

nesting.info2 <- nesting.info
nesting.info2$Day <- as.numeric(nesting.info2$Day)
AIC_mod.gls.DayNum <- gls(as.formula(paste("probitPlatform~",paste(model_var_gls,collapse="+"))),data=nesting.info2,correlation = nlme::corARMA(c(0.7762,0.0972),form=~1|Day/Tone,p=1,q=1),method = "ML")
summary(AIC_mod.gls)
```

Predictions
```{r}
pred.data <- as.data.frame(Test_Data.gls)
pred.var <- names(AIC_mod.gls$coefficients)[-c(1:52)]
pred.var <- c("Day","Tone","Bin",pred.var)
pred.data <- pred.data[,pred.var]
pred.data$Day <- pred.data$Day
pred.data$Tone <- as.factor(pred.data$Tone)
pred.data$Bin <- as.factor(pred.data$Bin)
pred.gls.mod <- predict(AIC_mod.gls.DayPoly,newdata = pred.data,na.action = NULL)
sum((as.numeric(Test_Data[,37]) - pred.gls.mod)^2)
```

Diagnostic Plots
```{r}
par(mfrow = c(1,2))
plot(AIC_mod.gls$fitted,AIC_mod.gls$residuals)
qqnorm(AIC_mod.gls$residuals)
abline(0,1)
```


Anova Test
```{r}
anova(AIC_mod.gls)
```

Comparison to TS
```{r}
plot(pred.gls.mod)
points(Test_Data.gls$probitPlatform,col = "red")
points(1:600,TS_predict_Day[["pred"]], col = "yellow")
```

