---
title: "Testing Model Format"
author: "Lukas"
date: "2023-01-30"
output: html_document
---

Note: First two tones of August 8th (day 2) are bad

Packages
```{r}
library(lme4)
library(VGAM)
library(car)
library(tseries)
library(forecast)
library(TSA)
library(doParallel)
library(parallel)
library(foreach)
```


Load MATLAB Data
```{r}
Aug_08 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_08_table.csv")
Aug_09 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_09_table.csv")
Aug_12 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_12_table.csv")
Aug_15 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Aug_15_table.csv")
Oct_12 <- read.csv("C:\\Users\\buhl5\\Desktop\\Plan B Files\\PMA17\\Oct_12_table.csv")
```

Remove Bad Tones from August 8th

```{r}
Aug_08 <- Aug_08[-c(1:60),]
```

Relabel Days
```{r}
Aug_08$Day <- 1
Aug_09$Day <- 2
Aug_12$Day <- 3
Aug_15$Day <- 4
Oct_12$Day <- 5
```



Day will likely be more useful when the days are all ran together with no gaps

Full Ephys Dataset
```{r}
Data <- rbind(Aug_08,Aug_09,Aug_12,Aug_15)
Test_Data <- Oct_12
rm(Aug_08,Aug_09,Aug_12,Aug_15,Oct_12)
```


Factors
```{r}
Data$Day <- as.factor(Data$Day)
Data$Tone <- as.factor(Data$Tone)
Data$Bin <- as.factor(Data$Bin)
```


Load Behavior Data
```{r}
## I Need to run August 5th through Analysis
Aug_08_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_08_new_R_TB_PMA17.csv")
Aug_09_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_09_new_R_TB_PMA17.csv")
Aug_12_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_12_new_R_TB_PMA17.csv")
Aug_15_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\08_15_new_R_TB_PMA17.csv")
Oct_12_Behav <- read.csv("C:\\Users\\buhl5\\Downloads\\10_12_new_R_TB_PMA17.csv")
```

Pull Time Specific Bins currently time 0 to 29. Will likely be adjusted to only tones

```{r}
Aug_08_Behav_p  <-  Aug_08_Behav[Aug_08_Behav[,2] %in% c(0:29),3]
Aug_09_Behav_p  <-  Aug_09_Behav[Aug_09_Behav[,2] %in% c(0:29),3]
Aug_12_Behav_p  <-  Aug_12_Behav[Aug_12_Behav[,2] %in% c(0:29),3]
Aug_15_Behav_p  <-  Aug_15_Behav[Aug_15_Behav[,2] %in% c(0:29),3]
Oct_12_Behav_p  <-  Oct_12_Behav[Oct_12_Behav[,2] %in% c(0:29),3]
```

All Behave

```{r}
#Remove same August Stuff
Aug_08_Behav_p <- Aug_08_Behav_p[-c(1:60)]
Behave <- c(Aug_08_Behav_p,Aug_09_Behav_p,Aug_12_Behav_p,Aug_15_Behav_p)

Test_Data$Platform <- Oct_12_Behav_p
Data$Platform <- Behave
```


Clean
```{r}
rm(Aug_08_Behav,Aug_09_Behav,Aug_12_Behav,Aug_15_Behav,Oct_12_Behav)
rm(Aug_08_Behav_p,Aug_09_Behav_p,Aug_12_Behav_p,Aug_15_Behav_p,Oct_12_Behav_p,Behave)
```

Data needs to be adjusted to accommodate 0 and 1

```{r}
#Playing it safe using a loop

offset <- 0.02


for(i in 1:length(Data$Platform)){
  
  if(Data$Platform[i] == 0){Data$Platform[i] <- offset}
  if(Data$Platform[i] == 1){Data$Platform[i] <- 1 - offset}
}

for(i in 1:length(Test_Data$Platform)){
  
  if(Test_Data$Platform[i] == 0){Test_Data$Platform[i] <- offset}
  if(Test_Data$Platform[i] == 1){Test_Data$Platform[i] <- 1 - offset}
}
```

Transform Response to make it fit regression requirement
```{r}
Data$probitPlatform <- probitlink(Data$Platform)
Test_Data$probitPlatform <- probitlink(Test_Data$Platform)
```



Is it autoregressive or moving average?

```{r}
plot(ts(Data$probitPlatform))
```


```{r}
pdf(file = "C:\\Users\\buhl5\\Desktop\\Plot PDFs\\TS_ACF.pdf",
    width = 5,
    height =3)
acf(ts(Data$probitPlatform),main="Autocorrelation of Probit Platform Time")
dev.off()
```

```{r}
pdf(file = "C:\\Users\\buhl5\\Desktop\\Plot PDFs\\TS_PACF.pdf",
    width = 5,
    height =3)
pacf(ts(Data$probitPlatform),main="Partial Autocorrelation of Probit Platform Time")
dev.off()
```
Suggests P should be 4 since it is the last significant lag before most remaining become not significant.





ADF test
```{r}
adf.test(Datats, alternative="stationary", k=0)
adf.test(Datats, alternative="stationary")
tseries::pp.test(Datats,alternative = "stationary",type = "Z(t_alpha)",lshort = F)
```

Lagged Differences
```{r}
acf(diff(Data$probitPlatform),main="Autocorrelation of Lagged Differences of Platform Time")
```

1,2,3,4,12,20,30 all appear significant especially 2,3,30


```{r}
Datats <- ts(Data$probitPlatform,frequency = 30,start = c(1,1))
```

Auto ARIMA for p,d,q

Previous graphs suggest that p is approximately 1-5 and q could be up to 30

```{r}
start <- Sys.time()
start

auto <- auto.arima(Datats, stationary = T,  seasonal = T, stepwise = T,ic = "aic",max.p = 4, max.q = 8, max.P = 3, max.Q = 3, max.order = 50,trace = T, test = "adf",approximation = T)
summary(auto)

end <- Sys.time()
end-start
```

Auto ARIMA suggests ARIMA (1,0,2)(2,0,0)[30] AIC 6667.31

Model Fit
```{r}
time.series.fit <- arima(Datats,order=c(1,0,2),
                 seasonal=list(order=c(2,0,0),period=30))
```


```{r}
plot(rstandard(time.series.fit))
hist(rstandard(time.series.fit))
qqnorm(rstandard(time.series.fit))
abline(0,1)
acf(time.series.fit$residuals)
tseries::adf.test(time.series.fit$residuals, alternative="stationary", k=0)
tseries::adf.test(time.series.fit$residuals, alternative="stationary")
car::dwt(as.numeric(time.series.fit$residuals),max.lag = 30)
(DWT = car::dwt(as.numeric(time.series.fit$residuals),max.lag = 30))
DWT.r <- as.data.frame(cbind(Lag = c(1:30), DWT = DWT))
```


